---
title: "Fire Visualizations"
date: "March 3, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
# fire time series
library(raster)
library(plyr)
library(dplyr)
library(tidyr)
library(broom)
library(ggplot2)
#library(doMC)
library(doParallel)
library(sf)
```


 
```{r echo=T, message=F, warning=F }

  # read in Regions 
  CC4a_reg = read_sf('./Boundries/CC4a_RegionsSub.shp')
  CC4a_reg  = st_transform(CC4a_reg, "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")
  # dissolve multipart Sierra Mountain feature
  # CC4a_reg = ms_dissolve(CC4a_reg,field = 'Region')
  
  # select SF 
  region_aoi = CC4a_reg$Region[7] 
  aoi = st_bbox(CC4a_reg[CC4a_reg$Region==region_aoi,])
  
  # read in data
  setwd('./bayarea/')
  l1r45 <- read.csv('./rcp45-low.csv',as.is=T)
  l2r45 <- read.csv('./rcp45-central.csv',as.is=T)
  l3r45 <- read.csv('./rcp45-high.csv',as.is=T)
  l1r85 <- read.csv('./rcp85-low.csv',as.is=T)
  l2r85 <- read.csv('./rcp85-central.csv',as.is=T)
  l3r85 <- read.csv('./rcp85-high.csv',as.is=T)
  
  # set up
  baf <- rbind(l1r45,l2r45,l3r45,l1r85,l2r85,l3r85)
  dim(baf)
  
  head(baf)
  baf$year <- as.numeric(substr(baf$date,4,7))
   
  #scenario groups
  baf$pop_recode = revalue(baf$population, c("L_mu"="Low", "bau_mu"="BAU",'H_mu'='High'))
  baf$scen_pop = paste(baf$scenario,baf$pop_recode,sep='_')
  baf$model_scen_pop = paste(baf$name, baf$scenario,baf$pop_recode,sep='_')
  baf$model = baf$name
  
  # reorder population factor
  baf$pop_recode = ordered(baf$pop_recode, levels = c("Low", "BAU", "High"))

```

# Cal adapt model agreement (westerling)
```{r}
  # download all from http://cal-adapt.org/data/
  
  # # unzip everything
  # setwd('C:/Users/mmann/Downloads/')
  # ziped = list.files('.',pattern='.zip',full.names = T)
  # for(file_zip in ziped){
  #   unzip(file_zip,exdir=tools::file_path_sans_ext(file_zip)) 
  # }
  # 
  # list.of.files <- list.files('C:/Users/mmann/Downloads/', ".tif$",recursive = T, full.names = T)
  # file.copy(list.of.files, to = 'C:/Users/mmann/Desktop/CA_Wildfire_CMIP5/Westerling_Fire_Runs/',overwrite = T)    

  decade_group_names = list(c(1976,2000), c(2001,2025), c(2026,2050),c(2051,2075),c(2076,2100))
  # model_abrev = c("CanESM2","CNRM-CM5","HadGEM2-ES","MIROC5" )
  # pop_scenario = c('bau') #,'L','H')
  # scenario = c(45,85)
   
  fire_tif = data.frame(fire_tif=list.files('./Westerling_Fire_Runs/',
                                            pattern='.tif', full.names = T),stringsAsFactors = F )

  # calculate average burned area for base year 
  # filter out which fire model runs to include in stacks 
  pop_scen = 'bau'
  clim_scen = '_85|45_'  # '_85|45_' to get both
  threshold = 0.75  # percentage that should agree eg 0.75

  # create mean base year 
  search_range = paste(seq(decade_group_names[[1]][1],decade_group_names[[1]][2]),collapse='|')
  mean_base_year = fire_tif %>% filter(grepl(search_range,fire_tif))%>% filter(grepl(pop_scen,fire_tif))%>% filter(grepl(clim_scen,fire_tif)) %>% .$fire_tif %>% stack() %>% mean(.,na.rm=T)
  
  # create stacks by decade group 
  search_range = paste(seq(decade_group_names[[2]][1],decade_group_names[[2]][2]),collapse='|')
  stack_2001_2025 = fire_tif %>% filter(grepl(search_range,fire_tif))%>% filter(grepl(pop_scen,fire_tif))%>% filter(grepl(clim_scen,fire_tif)) %>% .$fire_tif %>% stack()  #.$fire_tif is column name of paths
  
  search_range = paste(seq(decade_group_names[[3]][1],decade_group_names[[3]][2]),collapse='|')
  stack_2026_2050 = fire_tif %>% filter(grepl(search_range,fire_tif))%>% filter(grepl(pop_scen,fire_tif))%>% filter(grepl(clim_scen,fire_tif))  %>% .$fire_tif %>% stack()  
  
  search_range = paste(seq(decade_group_names[[4]][1],decade_group_names[[4]][2]),collapse='|')
  stack_2051_2075 = fire_tif %>% filter(grepl(search_range,fire_tif))%>% filter(grepl(pop_scen,fire_tif))%>% filter(grepl(clim_scen,fire_tif))  %>% .$fire_tif %>% stack()  
  
  search_range = paste(seq(decade_group_names[[5]][1],decade_group_names[[5]][2]),collapse='|')
  stack_2076_2100 = fire_tif %>% filter(grepl(search_range,fire_tif))%>% filter(grepl(pop_scen,fire_tif))%>% filter(grepl(clim_scen,fire_tif))  %>% .$fire_tif %>% stack()  
  
  
  # calculate mean difference for each period and model
  for(stack_name in ls(envir=.GlobalEnv, pattern="stack_")){
    st = stack(get(stack_name))
    #iterate through models
    for(model in unique(sapply(strsplit(names(st),"\\."),function(x){x[[1]][1]})) ){
        subset = st[[which(grepl(model,names(st)))]] %>% mean(.,na.rm=T)
        assign(paste('model_mean',model,stack_name,sep='_'),subset)
    }
  }
   
  # calc differences between future and base year mean
  for(stack_name in ls(envir=.GlobalEnv, pattern="model_mean_")){
    st = stack(get(stack_name))
    assign(paste('dif',stack_name,sep='_'),(st-mean_base_year))
  }
   
  
  # convert each model run to 1 for + changes and -1 for - changes
  for(dif_name in ls(envir=.GlobalEnv, pattern="dif_model")){
    dif = get(dif_name)#stack(get(dif_name))
    dif[dif>0]  =  1  # more fires
    dif[dif==0] =  0
    dif[dif<0]  = -1 # less fires 
    assign(dif_name,dif)
  }
   
  
   # create dif_model_mean stacks by decade group 
  search_range = paste(seq(decade_group_names[[2]][1],decade_group_names[[2]][2]),collapse='|')
  stack_2001_2025_names = ls(pattern = 'dif_model_mean') %>% as.data.frame(stringsAsFactors = F) %>% filter(grepl(search_range,.))%>% .$.   
  stack_2001_2025 = stack(mget(stack_2001_2025_names))   
  
  search_range = paste(seq(decade_group_names[[3]][1],decade_group_names[[3]][2]),collapse='|')
  stack_2026_2050_names = ls(pattern = 'dif_model_mean') %>% as.data.frame(stringsAsFactors = F) %>% filter(grepl(search_range,.))%>% .$.   
  stack_2026_2050 = stack(mget(stack_2026_2050_names))  
  
  search_range = paste(seq(decade_group_names[[4]][1],decade_group_names[[4]][2]),collapse='|')
  stack_2051_2075_names = ls(pattern = 'dif_model_mean') %>% as.data.frame(stringsAsFactors = F) %>% filter(grepl(search_range,.))%>% .$.   
  stack_2051_2075 = stack(mget(stack_2051_2075_names))  
   
  search_range = paste(seq(decade_group_names[[5]][1],decade_group_names[[5]][2]),collapse='|')
  stack_2076_2100_names = ls(pattern = 'dif_model_mean') %>% as.data.frame(stringsAsFactors = F) %>% filter(grepl(search_range,.))%>% .$.   
  stack_2076_2100 = stack(mget(stack_2076_2100_names))  
    
  # clean up 
  rm(list=c('stack_2001_2025_names','stack_2026_2050_names','stack_2051_2075_names','stack_2076_2100_names','stack_name'))
  rm(list=ls(pattern = 'model_mean_fire'))
  
  # find models that agree by %
  threshold_n = round( dim(get( "stack_2001_2025"))[3]*threshold) # num that must agree
  for(dif_name in ls(envir=.GlobalEnv, pattern="^stack_")){   #^ is "starts with"
    print(dif_name)
    sum_s =  sum(stack(get(dif_name))) # get sum of +1 -1 fires
    print(sum_s)
    agg_s = sum_s
    agg_s[!is.na(agg_s)] = 0  # get empty raster
    agg_s[sum_s>=threshold_n]  = 1  # agree +
    agg_s[sum_s<=(-1*threshold_n)]  = -1  # agree -
    assign(paste('agree',dif_name,sep='_'),agg_s)
  }
 
  # plot(agree_stack_2001_2025,main=paste('agree_',threshold,'perct_2001_2025',sep=''))
  # plot(agree_dif_2026_2050,main=paste('agree_',threshold,'perct_2026_2050',sep=''))
  # plot(agree_dif_2051_2075,main=paste('agree_',threshold,'perct_2051_2075',sep=''))
  # plot(agree_dif_2076_2100,main=paste('agree_',threshold,'perct_2076_2100',sep=''))

  # convert to data frame and set MFRI labels 
  for(dif_name in ls(envir=.GlobalEnv, pattern="agree_stack_")){
      df = data.frame(rasterToPoints(get(dif_name)))
      if(length(unique(df[,3]))==3){
          df$MFRI =   revalue(as.character(df[,3]), c('-1'='+ MFRI', '0'='None','1'='- MFRI'))  
          # +1 is more fire -MFRI is more fire
          df$MFRI = ordered(df$MFRI, levels =c('+ MFRI','None','- MFRI'))
      } 
      if(length(unique(df[,3]))==2){
          df$MFRI =   revalue(as.character(df[,3]), c('-1'='+ MFRI', '0'='None' ))  
          # 1 is more fire
          df$MFRI = ordered(df$MFRI, levels =c('+ MFRI','None'))
      }
      
      assign(paste('df',dif_name,sep='_'),df)
  }
  
  
  # plot agreement  
  ggplot() + geom_raster(data=na.omit(df_agree_stack_2001_2025), aes(x=x, y=y, fill=factor(MFRI)))  + geom_sf(data=CC4a_reg,colour = "grey30", fill = NA,size=.75) +   coord_sf( xlim=c(aoi[1],aoi[3]),ylim=c(aoi[2],aoi[4]) )+ scale_fill_manual(values = c( "#31B404","grey","#DF013A" ),name="Agreement")  + labs(fill='Human Effect\non Fire Count') +ggtitle(paste('2001_2025',clim_scen,threshold))
  ggsave(paste(pop_scen,gsub('\\|','',clim_scen),'agree',threshold,paste('2001_2025_Westerling_agreement.png'),sep='_'))
  
  ggplot() + geom_raster(data=na.omit(df_agree_stack_2026_2050), aes(x=x, y=y, fill=factor(MFRI)))  + geom_sf(data=CC4a_reg,colour = "grey30", fill = NA,size=.75) +   coord_sf( xlim=c(aoi[1],aoi[3]),ylim=c(aoi[2],aoi[4]) )+ scale_fill_manual(values = c("#31B404","grey","#DF013A"),name="Agreement")  + labs(fill='Human Effect\non Fire Count') +ggtitle(paste('2026_2050',clim_scen,threshold))
  ggsave(paste(pop_scen,gsub('\\|','',clim_scen),'agree',threshold,'2026_2050_Westerling_agreement.png',sep='_'))

  ggplot() + geom_raster(data=na.omit(df_agree_stack_2051_2075), aes(x=x, y=y, fill=factor(MFRI)))  + geom_sf(data=CC4a_reg,colour = "grey30", fill = NA,size=.75) +   coord_sf( xlim=c(aoi[1],aoi[3]),ylim=c(aoi[2],aoi[4]) )+ scale_fill_manual(values = c("#31B404","grey","#DF013A"),name="Agreement")  + labs(fill='Human Effect\non Fire Count') +ggtitle(paste('2051_2075',clim_scen,threshold))
  ggsave(paste(pop_scen,gsub('\\|','',clim_scen),'agree',threshold,'2051_2075_Westerling_agreement.png',sep='_'))

    ggplot() + geom_raster(data=na.omit(df_agree_stack_2076_2100), aes(x=x, y=y, fill=factor(MFRI)))  + geom_sf(data=CC4a_reg,colour = "grey30", fill = NA,size=.75) +   coord_sf( xlim=c(aoi[1],aoi[3]),ylim=c(aoi[2],aoi[4]) )+ scale_fill_manual(values = c("#31B404","grey","#DF013A"),name="Agreement")  + labs(fill='Human Effect\non Fire Count') +ggtitle(paste('2076_2100',clim_scen,threshold))
   ggsave(paste(pop_scen,gsub('\\|','',clim_scen),'agree',threshold,'2076_2100_Westerling_agreement.png',sep='_'))

```

```{r}

  # # calc differences between future and base year mean
  # diff_stack = function(stack_in, sub_in, nworkers=5){
  #    cl <- makeCluster(nworkers)
  #    registerDoParallel(cl)
  #    ras_back = foreach(layeri=1:dim(stack_in)[3], .inorder=T,.packages=c('raster')) %dopar% {
  #          return(stack_in[[layeri]]-sub_in)
  #      }
  #     stopCluster(cl)
  #     out_stack = stack(ras_back)
  #     names(out_stack) = sapply(strsplit(names(stack_in),"\\."),function(x){x[[1]][1]}) # remove year from name
  #     return(out_stack)
  # }
  # 
  # dif_2001_2025 = diff_stack(stack_2001_2025,mean_base_year)
  # dif_2026_2050 = diff_stack(stack_2026_2050,mean_base_year)
  # dif_2051_2075 = diff_stack(stack_2051_2075,mean_base_year)
  # dif_2076_2100 = diff_stack(stack_2076_2100,mean_base_year)
  
```



\newpage
## plot by climate and population scenario
```{r echo=T, message=F, warning=F }
# plot by climate and population scenario
ggplot(data=baf, aes(x=year,y=value,group=scen_pop,color=scen_pop))+geom_point(alpha=0.3)+
  geom_smooth(alpha=.5)+geom_smooth(se=F)+ ylab('Area Burned')+xlab('Year') + 
  labs(color='Climate & Population\nScenario')+coord_cartesian(ylim=c(5000,19000))  
```

\newpage
## plot lm models by groups
```{r echo=T, message=F, warning=F }
# plot lm models by groups
ggplot(data=baf, aes(x=year,y=value,group=model_scen_pop,color=model_scen_pop))+
  geom_point(alpha=0.3)+geom_smooth(se = FALSE, method = "lm")+ylab('Area Burned')+
  xlab('Year') + facet_wrap(~model)+labs(color='Climate & Population\nScenario')+  
  coord_cartesian(ylim=c(5000,19000))+theme(legend.position="none")
```
\newpage
```{r echo=T, message=F, warning=F }
ggplot(data=baf, aes(x=year,y=value,group=model_scen_pop,color=model_scen_pop))+
  geom_point(alpha=0.3)+geom_smooth(se = FALSE, method = "lm")+ylab('Area Burned')+
  xlab('Year') + facet_wrap(~pop_recode)+labs(color='Climate & Population\nScenario')+
  coord_cartesian(ylim=c(5000,19000))+theme(legend.position="none")
```
\newpage
```{r echo=T, message=F, warning=F }
ggplot(data=baf, aes(x=year,y=value,group=model_scen_pop,color=model_scen_pop))+
  geom_point(alpha=0.3)+geom_smooth(se = FALSE, method = "lm")+ylab('Area Burned')+
  xlab('Year') + facet_wrap(~scenario)+labs(color='Climate & Population\nScenario')+
  coord_cartesian(ylim=c(5000,19000))+theme(legend.position="none")
```
\newpage

##Calculate regression coefficients
```{r echo=T, message=F, warning=F }
# get regression coefficients
baf_reg = baf %>% group_by(model,scenario, pop_recode) %>% do(fit = lm(value~year, data = .))
baf_coef = tidy(baf_reg, fit) %>% as.data.frame()
head(baf_coef)

# select signficant coefficients
baf_sig = baf_coef  %>% filter(term == "year" & p.value < 0.05) %>% 
  select(model,scenario, pop_recode,estimate )
head(baf_sig)
```
 

##Plot distribution of regression coefficients
```{r echo=T, message=F, warning=F }
ggplot(data = baf_sig)+geom_boxplot(aes(x=model,y=estimate,fill=model))+ 
  labs(fill='Climate\nModel') +ylab(expression(Delta~' Burned Area / Year') )+
  xlab('Climate Model')
```
\newpage
```{r echo=T, message=F, warning=F }
ggplot(data = baf_sig)+geom_boxplot(aes(x=scenario,y=estimate,fill=scenario))+ 
  labs(fill='Climate\nScenario') +ylab(expression(Delta~' Burned Area / Year') )+
  xlab('Climate Model')
```
\newpage
```{r echo=T, message=F, warning=F }
# fill by climate and population scenaario
ggplot(data = baf_sig)+geom_boxplot(aes(x=model,y=estimate, fill= scenario))+ 
  labs(fill='Climate\nScenario') +ylab(expression(Delta~' Burned Area / Year') )+
  xlab('Climate Model')
```
\newpage
```{r echo=T, message=F, warning=F }
ggplot(data = baf_sig)+geom_boxplot(aes(x=model,y=estimate, fill= pop_recode)) + 
  labs(fill='Population\nScenario')+ylab(expression(Delta~' Burned Area / Year') )+
  xlab('Climate Model')
```
\newpage
```{r echo=T, message=F, warning=F }
# other versions
ggplot(data = baf_sig)+geom_boxplot(aes(x=scenario,y=estimate,fill=model))+ 
  labs(fill='Climate\nScenario') +ylab(expression(Delta~' Burned Area / Year') )+
  xlab('Climate Model')
```
\newpage
```{r echo=T, message=F, warning=F }
ggplot(data = baf_sig)+geom_boxplot(aes(x=scenario,y=estimate,fill=pop_recode))+ 
  labs(fill='Climate\nScenario') + ylab(expression(Delta~' Burned Area / Year') )+
  xlab('Climate Model')

```
 
\newpage

#Calculate maximum burn values
```{r echo=T, message=F, warning=F }
# get regression coefficients
baf_max = baf %>% group_by(model,scenario, pop_recode) %>%   summarize(max = max(value, na.rm = T)) %>% 
  as.data.frame()
head(baf_max)
```


##Plot max burned area
```{r echo=T, message=F, warning=F }
ggplot(data = baf_max)+geom_boxplot(aes(x=model,y=max,fill=model))+ 
  labs(fill='Climate\nModel') +ylab('Maximum Burned Area' )+
  xlab('Climate Model')
```
\newpage
```{r echo=T, message=F, warning=F }
ggplot(data = baf_max)+geom_boxplot(aes(x=scenario,y=max,fill=scenario))+ 
  labs(fill='Climate\nScenario') +ylab('Maximum Burned Area')+
  xlab('Climate Model')
```
\newpage
```{r echo=T, message=F, warning=F }
# fill by climate and population scenaario
ggplot(data = baf_max)+geom_boxplot(aes(x=model,y=max, fill= scenario))+ 
  labs(fill='Climate\nScenario') +ylab('Maximum Burned Area' )+
  xlab('Climate Model')
```
\newpage
```{r echo=T, message=F, warning=F }
ggplot(data = baf_max)+geom_boxplot(aes(x=model,y=max, fill= pop_recode)) + 
  labs(fill='Population\nScenario')+ylab('Maximum Burned Area' )+
  xlab('Climate Model')
```
\newpage
```{r echo=T, message=F, warning=F }
# other versions
ggplot(data = baf_max)+geom_boxplot(aes(x=scenario,y=max,fill=model))+ 
  labs(fill='Climate\nScenario') +ylab('Maximum Burned Area'  )+
  xlab('Climate Model')
```
\newpage
```{r echo=T, message=F, warning=F }
ggplot(data = baf_max)+geom_boxplot(aes(x=scenario,y=max,fill=pop_recode))+ 
  labs(fill='Climate\nScenario') + ylab('Maximum Burned Area')+
  xlab('Climate Model')

```


